\chapter{Background}

This chapter provides some background knowledge about the facts and technology that support this project and make it possible it this point of time. It also informs why it is getting more and more important to the software industry.

\section{Mobile Device Sensors}
Over the last decades the evolution mobile devices began with a wireless telephone far away from pocket size. Over the years the mobile devices got displays, SMS, telephone books, games and a lot more. In 2007, Steve Jobs introduced the first iPhone and with it the age of the smartphone  \cite{laugesen2010factors}. Over the years, smart-phones became pocket size computers with a better display resolution than the most televisions and the computing power of what desktop pc users could just dream about a few years ago. More and more sensors were packet into the small handy devices were made easy accessible by developers.\\
The sensors range from proximity detection over accelerometer to humidity sensors etc. Google even engineered a system for 3D objects and indoor environments with just a single device in real time \cite{schops20153d}.\\
So, a lot of people own the hardware with the capability to collect rich context information and they even carry it with them all the time and could be used for support and improve the peoples work and environment.

\section{The importance of Software Metrics}
Software is becoming more complex than ever and used in almost every environment. The deadlines in professional software projects are very strict and there is no time to develop everything from scratch. Development teams depend on libraries rather than reinventing the wheel over and over again. \\
The problem is that it's so easy to start programming and learning everything online from various sources. A lot of people do programming for a hobby and the quality and performance doesn't matter to them as much as in a professional environment. They also create libraries and frameworks with their quality standards. Also the increasing amount of open source libraries and the dependencies they create makes it very hard but also very important to professionals that they can trust these libraries and the quality. At this point of time the only indicators are user ratings and the amount of times it's been used in different frameworks and projects. Some frameworks are also recommended in public reviews articles. \\
Frameworks and libraries also need to be more dynamic and maintained. Operating systems and programming languages are being updated more frequently which requires fast changes. A constant measuring of quality could at the first place give direct feedback that the quality stays constant after changes and as well helps the developer to create a better structure and code to improve the maintainability at the first place. \\
The most used platform in the open source community is Github. Github is based on git and is a web server that can be used to host software projects and allow to make them accessible to others developers  \cite{dabbish2012social}.